An Invariant Framework for Conducting Reproducible Computational Science

### 2015-03-30 intro.tex
Proof-reading

Reproducibility is a cornerstone of the scientific method~\cite{borgman2012data}. 
Its importance is underscored by its ability to advance science---reproducing by verifying and validating a scientific result leads to improved understanding, thus increasing possibilities of reusing or extending the result. 
Ensuring reproducibility of a scientific result, however, often entails detailed documentation and specification of the involved scientific method. Historically, this has been achieved through text and proofs in a publication. 
As computation pervades the sciences and transforms the scientific method, mere text is no longer sufficient. 
In particular, apart from textual descriptions describing the result, a reproducible result must also include several computational artifacts, such as software, data,  environment variables, and state of computation that are involved in the adopted scientific method \cite{Sole}.

Virtualization has emerged as a promising approach for reproducing computational scientific results. One approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated, and executable record of computations, especially computations whose results are destined for publication and/or re-use; images, like files, can then be shared \cite{Lampoudi}. 
However, the resulting image may be too big to share or distribute widely. An alternative light-weight form of virtualization is to encapsulate only the application software, along with all its necessary dependencies, into a self-contained package. The encapsulation is achieved by OS-level sandboxing techniques that interpose application system calls, and copy the necessary dependencies (data, libraries, code, etc.) into a package, making it lighter weight than a VMI~\cite{guo2011cde}. However, the package is not longer an executable record of the computation and still needs an accompanying operating system for execution.

While both approaches provide mechanisms for encapsulating the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result. In general, in the absence of reproducible policy guidelines, such guarantees can be hard to provide. However, we can certainly improve upon the guarantees by \emph{preserving} the encapsulated computations in such a way that they are always reproducible. A preservation mechanism, if present, can make it easy to install images or packages, alter implicit dependencies that are part of computation as software components evolve or become deprecated, and finally provide  mechanisms for documentation that make computations easy to understand.

@@

There are two approaches to address the preservation challenge. By either introducing tools that help document dependencies and provide software attribution within VMIs or packages, or alternatively by using software delivery mechanisms, such as, centralized package management, Linux containers, and the, more recent, Docker framework. We examined the first approach previously in \cite{SoftProv}. 
In this paper we examine the second approach.  
In particular, we consider the light-weight virtualization approaches, because, we believe that a combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers.  
A package created by those light-weight approaches encapsulates all the necessary dependencies of an application, and can be used to repeat the application through different sandbox mechanisms, including Parrot~\cite{thain2005parrot}, CDE, PTU~\cite{PTU}, chroot, and Docker~\cite{boettiger2015introduction}.

We do not claim our solution is the only way to preserve applications. Broadly, there are two different approaches to preserve applications: force cleanliness or measure the mess. The first approach forces users to specify the execution environment for an application in a well organized way. In the latter one end users construct the environment as desired, and the complexity of the environment is measured in terms of its dependencies. 
Our objective here is to measure the mess as-is and then preserve it over time.

To conduct a thorough examination, we consider real-world complex high energy physics (HEP) applications, independently developed by two groups, that must be reproduced so that the entire HEP community can benefit from the analysis. We describe challenges in reproducing the applications, and consider the extent to which reproducibility requirements can be satisfied with light-weight virtualization approaches and software delivery mechanisms. We propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms for efficiently capturing, invariantly preserving and practically deploying applications.
We measure the performance overhead of  light-weight virtualization and software delivery approaches, and show the preserved packages can be distributed to allow reproduction and verification.

>>>>

Reproducibility is a cornerstone of the scientific method~\cite{borgman2012data}. 
Its ability to advance science underscores its importance---reproducing by verifying and validating a scientific result leads to improved understanding, thus increasing possibilities of reusing or extending the result. 
Ensuring the reproducibility of a scientific result, however, often entails detailed documentation and specification of the involved scientific method. Historically, text and proofs in a publication have achieved this end. 
As computation pervades the sciences and transforms the scientific method, mere text is no longer sufficient. 
In particular, apart from textual descriptions describing the result, a reproducible result must also include several computational artifacts, such as software, data,  environment variables, and the state of computation that are involved in the adopted scientific method \cite{Sole}.

Virtualization has emerged as a promising way to reproduce computational scientific results. One such approach is to conduct the entire computation relating to a scientific result within a virtual machine image, and then share the resulting image. This way VMIs become an authoritative, encapsulated, and executable record of computations, especially computations whose results are destined for publication and/or re-use; images, like files, can then be shared \cite{Lampoudi}. 
The resulting image, however, may be too big to share or distribute widely. An alternative light-weight form of virtualization is to encapsulate only the application software along with all its necessary dependencies into a self-contained package. The encapsulation is achieved by OS-level sandboxing techniques that interpose application system calls and copy the necessary dependencies (data, libraries, code, etc.) into a package, making it lighter weight than a VMI~\cite{guo2011cde}. Yet, the package is not longer an executable record of the computation and still requires an accompanying operating system for execution.

While both approaches provide mechanisms for encapsulating the computations associated with a scientific result, neither form of virtualization provides any guarantee that the included pieces of software will indeed reproduce the associated scientific result. In general, in the absence of reproducible policy guidelines, such guarantees can be difficult to provide. Preserving the encapsulated computations in such a way that they are always reproducible will improve upon the guarantees. A preservation mechanism can increase the ease of image or package installation, alter dependencies implicit to computation as software components evolve or become deprecated, and provide mechanisms for documentation that make computations easy to understand.

@@

The two approaches that address the preservation challenge are as follows: one, the introduction of tools that help document dependencies and provide software attribution within VMIs or packages; and two, the use of software delivery mechanisms such as centralized package management, Linux containers, and the more recent Docker framework. We examined the first approach previously in \cite{SoftProv}. 
In this paper we examine the second approach.  We consider in particular the lightweight virtualization because we believe together with more standardized software delivery mechanisms, the two combined can address the reproducibility challenge for a wide variety of scientific researchers. A package created by those lightweight approaches encapsulates all the necessary dependencies of an application, and can be used to repeat the application through different sandbox mechanisms, including Parrot
%In particular, we consider the light-weight virtualization approaches, because, we believe that a combination of light-weight approaches with more standardized software delivery mechanisms can lead to addressing the reproducibility challenge for a wide variety of scientific researchers.  
%A package created by those light-weight approaches encapsulates all the necessary dependencies of an application, and can be used to repeat the application through different sandbox mechanisms, including Parrot
~\cite{thain2005parrot}, CDE~\cite{Guo}, PTU~\cite{PTU}, chroot, and Docker~\cite{boettiger2015introduction}.

Of course our solution represents only one way to preserve applications. Broadly, two different approaches to preserve applications have been adopted: force cleanliness or measure the mess. The former forces users to specify the execution environment for an application in a well-organized way. The latter causes end users to construct the environment as desired, and the complexity of the environment is measured in terms of its dependencies. Our objective here is to measure the mess as-is and then preserve it over time.

%We do not claim our solution is the only way to preserve applications. Broadly, there are two different approaches to preserve applications: force cleanliness or measure the mess. The first approach forces users to specify the execution environment for an application in a well organized way. In the latter one end users construct the environment as desired, and the complexity of the environment is measured in terms of its dependencies. 
%Our objective here is to measure the mess as-is and then preserve it over time.

To conduct a thorough examination, we consider real-world complex high energy physics (HEP) applications, independently developed by two groups, that must be reproduced so that the entire HEP community can benefit from the analysis. We describe challenges faced in reproducing the applications, and we consider the extent to which reproducibility requirements can be satisfied with lightweight virtualization approaches and software delivery mechanisms. We propose an invariant framework for computational reproducibility that combines lightweight virtualization with software delivery mechanisms for efficiently capturing, invariantly preserving, and practically deploying applications. We measure the performance overhead of lightweight virtualization and software delivery approaches, and show how the preserved packages can be distributed to allow reproduction and verification.
%To conduct a thorough examination, we consider real-world complex high energy physics (HEP) applications, independently developed by two groups, that must be reproduced so that the entire HEP community can benefit from the analysis. We describe challenges in reproducing the applications, and consider the extent to which reproducibility requirements can be satisfied with light-weight virtualization approaches and software delivery mechanisms. We propose an invariant framework for computational reproducibility that combines light-weight virtualization with software delivery mechanisms for efficiently capturing, invariantly preserving and practically deploying applications.
%We measure the performance overhead of  light-weight virtualization and software delivery approaches, and show the preserved packages can be distributed to allow reproduction and verification.

### 2015-03-30 main.tex
Rename paper.tex to main.tex Use DOI instread of the URL for the DOI
Proof-reading

\begin{abstract}
\it Computational reproducibility depends on being able to isolate necessary and sufficient computational artifacts and preserve them for later re-execution.
Both isolation and preservation of artifacts can be challenging due to the complexity
of existing software and systems and the resulting implicit dependencies, resource distribution, and shifting compatibility of systems as time progresses---all conspiring
to break the reproducibility of an application. Sandboxing is a technique
that has been used extensively in OS environments for isolation of computational artifacts.
Several tools were proposed recently that employ sandboxing as a mechanism to ensure reproducibility.
However, none of these tools preserve the sandboxed application for re-distribution
to a larger scientific community---aspects that are equally crucial for ensuring reproducibility as sandboxing itself.
In this paper, we describe a combined sandboxing and preservation framework, which is efficient, invariant and
practical for large-scale reproducibility. We present case studies of complex high energy
physics applications and show how the framework can be useful for sandboxing, preserving and distributing applications.
We report on the completeness, performance,
and efficiency of the framework, and suggest possible standardization approaches.
\end{abstract}

\input {intro}
\input {tauroast}
\input {observation}
\input {evolution}
\input {measure}
\input {evaluation}
\input {related_work}
\section{Conclusions and Future work}
In this paper, we propose an invariant framework for conducting reproducible computational science - using light-weight virtualization approaches to preserve applications in the format of self-contained packages and using standardized software delivery mechanisms to deliver and distribute preserved packages.
We use two complex high energy physics applications to illustrate how the framework can help the original authors preserve and distribute the applications, and others reproduce the applications.

This paper focuses on how to measure the mess and track the used dependencies to preserve an application.
In the following work, we plan to explore how to preserve an application in an organized style - specifying the execution environment clearly.
How to preserve and improve the availability of remote network resources is another important problem to be explored.

For more details about how to preserve, distribute and repeat an application:

\indent \indent \indent \indent \indent \indent \indent \url{http://dx.doi.org/doi:10.7274/R0C24TCG}

@@

\bibliographystyle{plain}
\bibliography{cclpapers,this,sole}

>>>>

\begin{abstract}
\it Computational reproducibility depends on the ability to not only isolate necessary and sufficient computational artifacts but also to preserve those artifacts for later re-execution. Both isolation and preservation present challenges in large part due to the complexity of existing software and systems as well as the implicit dependencies, resource distribution, and shifting compatibility of systems that result over time---all of which conspire to break the reproducibility of an application. Sandboxing is a technique that has been used extensively in OS environments in order to isolate computational artifacts. Several tools were proposed recently that employ sandboxing as a mechanism to ensure reproducibility. However, none of these tools preserve the sandboxed application for re-distribution to a larger scientific community—aspects that are equally crucial for ensuring reproducibility as sandboxing itself. In this paper, we describe a framework of combined sandboxing and preservation, which is not only efficient and invariant, but also practical for large-scale reproducibility. We present case studies of complex high-energy physics applications and show how the framework can be useful for sandboxing, preserving, and distributing applications. We report on the completeness, performance, and efficiency of the framework, and suggest possible standardization approaches.

\end{abstract}

\vspace{-10pt}
\input {intro}
\vspace{-10pt}
\input {tauroast}
\vspace{-10pt}
\input {observation}
\vspace{-10pt}
\input {evolution}
\vspace{-10pt}
\input {measure}
\vspace{-10pt}
\input {evaluation}
\vspace{-10pt}
\input {related_work}
\vspace{-10pt}
\section{Conclusions and Future work}
In this paper, we propose an invariant framework for conducting reproducible computational science - using light-weight virtualization approaches to preserve applications in the format of self-contained packages and using standardized software delivery mechanisms to deliver and distribute preserved packages.
We use two complex high energy physics applications to illustrate how the framework can help the original authors preserve and distribute the applications, and others reproduce the applications.
This paper focuses on how to measure the mess and track the used dependencies to preserve an application.
In the following work, we plan to explore how to preserve an application in an organized style - specifying the execution environment clearly.
How to preserve and improve the availability of remote network resources is another important problem to be explored.

\small
\vspace{-10pt}
DOI for the experiment involved in the paper: \url{doi:10.7274/R0C24TCG}

@@

\vspace{-10pt}
\bibliographystyle{plain}
\bibliography{cclpapers,this,sole}

### 2015-03-30 observation.tex
Proof-reading

The \emph{TauRoast} and \emph{Athena} application specification was provided to us in the form of an
email which described, in prose, how to obtain the source,
build the program, and run it correctly on one specific
machine at our home institution, with no particular guarantee that
it will run anywhere else in the world. This minimal level of documentation about software is routine in the scientific world. 
We provide the challenges we faced in capturing the application details in a reproducible form and then preserving it for subsequent reuse:

@@

\item {\bf Identifying all dependencies.}  Due to the distributed nature of HEP applications, these applications depend on a large number of external and local dependencies.
External dependencies are often explicitly stated, such as when the application makes connections to Github resources, and CVS servers for downloading source. 
However, when the application has distributed execution then implicit network connections are present requiring us to identify dependencies on all machines where execution takes place. 
Implicit local dependencies arise due to mounted filesystems. In \emph{TauRoast}, the application data and code is distributed on five networked filesystems, and in \emph{Athena} on two networked filesystems. 
Since these filesystems appear local to the application machine, it is important to check and capture mounted filesystems and their respective mount points.

\item {\bf Configuration Complexity.} To correctly reproduce an application implies that run-time configurations and consistency checks on the available software are effectively captured and preserved. 
For CMS,  the {\tt scram} software management tool is used to locate
the appropriate version of software,  set environment variables such as the PATH, run any
tool-specific configuration, and do the same for all software on which it depends. A reproducible framework must capture the work of such software management tools so that they can conduct similar
checks on a new machine.

\item {\bf High Selectivity.}  
Although the total size of the resources accessed by HEP
programs is very large, the size of the data and software actually used are much smaller.
Often, an entire repository or data source is named within the script, but the program
only needs a handful of items from that source.  For example, the data is stored on an
HDFS filesystem with 11.6TB of data, but only 18GB are actually consumed by the program. 
Thus there is a size tradeoff of capturing dependencies mentioned in the program and dependencies actually used in the program.
A reproducible framework must include robust rules about not including superfluous dependencies, but including unused dependencies that may potentially get used during program execution.

\item {\bf Rapid Changes in Dependencies.}  Over the course of three months
between collecting the initial email, analyzing the program, and writing this
paper, the computing environment continuously changed.  The CMSSW software
distribution released a new version, the target execution environment was upgraded
to a new operating system, and the application switched from CVS to Git for obtaining
the software. In Athena, the computing environment can potentially change daily, since upgrades to the software framework occur on a nightly basis.  
While the users of this software seem be accustomed to constant change,
any preservation technique will have to be very cautious about relying upon an
external service, even one that may appear to be highly stable.

\item {\bf Dependencies for Reproducible Execution.} Capturing the necessary and sufficient dependencies that are part of an application is sufficient for repeatability, but possibly for not reproducibility.
Repeatability implies if a result depends on running program X, we must be able to run exactly X again. In reproducibility the goal is rarely limited to running
\emph{precisely} what a predecessor did. Often, the objective is to
change a parameter or a data input in order to see how the result is affected. To that end, the preservation system must capture enough of the surrounding
material to permit modifications to succeed. 
Further, a better understanding of how end users will consume preserved software will help to shape how
software should be preserved.
\end{itemize}

>>>>

The application specifications of \emph{TauRoast} and \emph{Athena} were provided to us in the form of an
email that described in prose how to obtain the source,
build the program, and run it correctly on one specific
machine at our home institution, but with no particular guarantee that
it will run anywhere else in the world. This minimal level of documentation about software is routine in the scientific world. 
Below we describe the challenges faced when capturing the application details in a reproducible form and then preserving them for subsequent reuse:

@@

\item {\bf Identifying all dependencies.}  Due to the distributed nature of HEP applications, these applications depend on a large number of external and local dependencies.
External dependencies are often explicitly stated, such as when the application makes connections to Github resources, and to CVS servers for downloading the source. 
When the application has distributed execution then implicit network connections may then be present, thus requiring us to identify dependencies on all machines where execution takes place. 
Implicit local dependencies arise as a result of mounted filesystems. In \emph{TauRoast}, the application data and code is distributed on five networked filesystems, and in \emph{Athena} on two networked filesystems. 
Since these filesystems appear local to the application machine, it is important to check and capture mounted filesystems and their respective mount points.

\item {\bf Configuration Complexity.} In order to correctly reproduce an application requires that run-time configurations and consistency checks on the available software are effectively captured and preserved. 
For CMS,  the {\tt scram} software management tool is used to locate
the appropriate version of software,  set environment variables such as the PATH, run any
tool-specific configuration, and do the same for all software on which it depends. A reproducible framework must capture the work of such software management tools so that the framework can conduct similar
checks on a new machine.

\item {\bf High Selectivity.}  
Although the total size of the resources accessed by HEP programs is very large, the size of the data and software actually used are much smaller. The script will often name an entire repository or data source, but the program needs only a handful of items from that source. For example, the data may be stored on an HDFS filesystem with 11.6TB of data, but the program may consume only 18GB. Thus there is a size tradeoff in terms of capturing dependencies mentioned in the program and dependencies actually used in the program. A reproducible framework must include robust rules about not including superfluous dependencies, but including unused dependencies that may potentially see much use during program execution.

\item {\bf Rapid Changes in Dependencies.}  
Over the course of three months between collecting the initial email, analyzing the program, and writing this paper, the computing environment saw continuous change. The CMSSW software distribution released a new version, the target execution environment was upgraded to a new operating system, and the application switched from CVS to Git for obtaining the software. In Athena, the computing environment has the potential for daily change since upgrades to the software framework occur on a nightly basis. While the users of this software seem accustomed to constant change, any new technique for preservation that may rely on an external service (every one that may appear highly stable) will require great caution.

\item {\bf Dependencies for Reproducible Execution.} Capturing the necessary and sufficient dependencies that are part of an application is sufficient for repeatability, but possibly for not reproducibility.
Repeatability implies that if a result depends on running program X, we must be able to run exactly X again. In reproducibility the goal is rarely limited to running
\emph{precisely} what a predecessor did. Often, the objective is to
change a parameter or a data input in order to see how the result is affected. To that end, the preservation system must capture enough of the surrounding
material in order to permit modifications to succeed. 
Further, a better understanding of how end users will consume preserved software will help to shape how
software should be preserved.
\end{itemize}

### 2015-03-30 tauroast.tex
Proof-reading

We study applications taken from two experiments of the CERN Large Hadron Collider, namely the ATLAS experiment and the CMS experiment. 
In LHC, the ATLAS and CMS experiments are distinct, 
developed independently by two entirely separate physics communities. Consequently, their applications  
have very different software distribution and data management frameworks, making it interesting to examine if common reproducibility frameworks and 
tools work across the two communities. 
One of the applications of the ATLAS experiment is the \emph{Athena} application, which is a general purpose processing framework including algorithms
for event reconstruction and data reduction \cite{calafiura2005athena}. The CMS experiment is conducted through an application termed  \emph{TauRoast}, which searches for specific 
cases where the Higgs boson decays to two tau leptons~\cite{chatrchyan2013search}.

Code and data in \emph{TauRoast} is available through five different networked filesystems which are mounted locally, an HDFS cluster for data, 
some configuration files were stored on CVMFS~\cite{blomer2011cernvm}, and a variety of software tools were on an NFS, PanFS and AFS systems.
In addition, code may exist in version control systems such as Git, CVS, and CMS Software Distribution (CMSSW). %~\cite{cms2006cms}. 
\begin{wrapfigure}{r}{0.4\textwidth}
\small
\centering
\includegraphics[width=.4\textwidth]{data-code-size.eps}
\caption{Inputs to Tau Roast}
\label{fig:data-code-size}
\end{wrapfigure}
Data that is input to \emph{TauRoast} is obtained by reducing it through a pipeline, as shown in Figure \ref{fig:data-code-size}. Consequently, the real input data may 
vary depending upon the science question being researched. Similarly the software may name many possible components but the used components are
smaller than the named ones.

In \emph{Athena} data is obtained through an external Dropbox-like system called the FaxBox, but does not go through any reduction steps. Code is obtained through 
CVMFS, which provides the analysis routines. However, depending upon the input data code and the configuration that will be invoked changes. 
Thus in \emph{Athena} the used code and configuration are dynamic depending upon input data, where as in \emph{TauRoast} the code and data are static, 
but the amount of data and code to include changes depending on the involved science.

>>>>

We study applications taken from two experiments of the CERN Large Hadron Collider, namely the ATLAS experiment and the CMS experiment. 
In LHC, the ATLAS and CMS experiments are distinct, 
developed independently by two entirely separate physics communities. Consequently, their applications  
have very different software distribution and data management frameworks, raising the question of whether common reproducibility frameworks and tools work across the two communities.  
%making it interesting to examine if common reproducibility frameworks and 
%tools work across the two communities. 
One of the applications of the ATLAS experiment is the \emph{Athena} application, which is a general purpose processing framework including algorithms
for event reconstruction and data reduction \cite{calafiura2005athena}. The CMS experiment is conducted through an application termed  \emph{TauRoast}, which searches for specific 
cases where the Higgs boson decays to two tau leptons~\cite{chatrchyan2013search}.

Code and data in \emph{TauRoast} are available through five different networked filesystems which are mounted locally, an HDFS cluster for data, 
some configuration files were stored on CVMFS~\cite{blomer2011cernvm}, and a variety of software tools were on an NFS, PanFS and AFS systems.
In addition, code may exist in version control systems such as Git, CVS, and CMS Software Distribution (CMSSW). %~\cite{cms2006cms}. 
\begin{wrapfigure}{r}{0.4\textwidth}
\small
\centering
\includegraphics[width=.4\textwidth]{data-code-size.eps}
\caption{Inputs to Tau Roast}
\label{fig:data-code-size}
\end{wrapfigure}
Data that is input to \emph{TauRoast} is obtained by reducing it through a pipeline, as shown in Figure \ref{fig:data-code-size}. Consequently, the real input data may 
vary depending upon the topic of research. Similarly the software may name many possible components but the used components are
smaller than the named ones.

Data in \emph{Athena} is obtained through an external Dropbox-like system called the FaxBox, but does not pass through any reduction steps. Code is obtained through 
CVMFS, which provides the analysis routines. 
The invoked configuration will change, however, depending upon the input data code.  Thus in Athena  the used code and configuration are dynamic depending upon input data, whereas in TauRoast the code and data are static, but the amount of data and code to include changes depending on the science involved.
